---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 Library statements ------------------------------------------------------

```{r}
library(tidyverse)
library(ggforce)
library(ggtext)
library(ggh4x)
library(seriation)
library(dynamicTreeCut)
library(colorspace)
```

# Order cities on population to get the more populous one when matched by name.

```{r}
cities <- maps::world.cities %>%
  filter(name != "Gold Coast" & country.etc != "Australia") %>% 
  # Because that is not the intended Gold Coast
  arrange(-pop)
```

# Place-country lookup table (LUT) for less populous areas and 
# other kinds of places (rivers, provinces, etc.).
# I didn't fact check this and the list is not complete.
# Mostly I googled these places and made an educated guess.
# Especially for African places that now have newer names than in the data this
# method is error prone.
```{r}
place_lut <- tribble(
  ~replace, ~by,
  "Havana", "Havanna",
  "Brest", "France", # Might have been confused with landlocked Brest, Belarus
  "Bahia", "Brazil",
  "Seville", "Sevilla",
  "Southeast Brazil", "Brazil",
  "Essequibo", "Guyana",
  "Cap Francais", "Cap-Haitien",
  "Rhode Island", "USA",
  "Bristol (RI)", "USA",
  "Zeeland", "Netherlands",
  "Hellevoetsluis", "Netherlands",
  "Pernambuco", "Brazil",
  "San Lucar", "Spain",
  "Honfleur", "France",
  "Dunkerque", "Dunkirk",
  "St. Thomas", "USA",
  "Kobenhavn", "Copenhagen",
  "Africa.", "Africa",
  "St. Kitts", "Saint Kitts and Nevis",
  "Demerara", "Guyana",
  "Cayes (Les)", "Les Cayes",
  "St. Vincent", "Saint Vincent and The Grenadines",
  "York River", "Richmond",
  "Maranhao", "Sao Luis",
  "Curacao", "Willemstad",
  "Saint John (Antigua)", "Antigua and Barbuda",
  "Para", "Brazil",
  "Upper James River", "USA",
  "Rappahannock", "USA",
  "Newcastle (Nevis)", "Saint Kitts and Nevis",
  "St. Croix", "US Virgin Islands",
  "St. Eustatius", "Netherlands Antilles",
  "Saint-Domingue", "Cap-Haitien",
  "Berbice", "Guyana",
  "Tortola", "British Virgin Islands",
  "Zion Hill", "Zions Hill",
  "St. Helena", "Saint Helena",
  "Maryland", "USA",
  "Patuxent", "USA",
  "Annapolis", "USA",
  "St. Lucia", "Saint Lucia",
  "South Potomac", "USA",
  "Ponta Negro", "Natal",
  "Petit-Goave", "Haiti",
  "Cape of Good Hope", "Cape Town",
  "Trinidad de Cuba", "Cuba",
  "Dutch Guianas", "Guyana",
  "Louisiana", "USA",
  "Cuba, south coast", "Cuba",
  "Taipu", "Brazil",
  "Falmouth (Jam)", "Jamaica",
  "Martha Brae", "Jamaica",
  "Ilha Grande", "Brazil",
  "Danish West Indies", "US Virgin Islands",
  "Catuamo and Maria Farinha", "Brazil",
  "Hispaniola", "Santo Domingo",
  "Hispanioloa, unspecified", "Santo Domingo",
  "South Carolina", "USA",
  "North Carolina", "USA",
  "Lower James River", "Richmond",
  # Data becomes more of a guess here
  "Whydah", "Benin",
  "Congo North", "Congo",
  "Malembo", "Angola",
  "Cape Lahou", "Grand-Lahou",
  "Gold Coast", "Ghana",
  "Ardra", "Benin",
  "Loango", "Congo",
  "Jacquin", "Benin", # https://www.slavevoyages.org/voyage/33330/map I think?
  "Boary", "Angola",
  "West Central Africa and St. Helena", "West Central Africa",
  "Badagry/Apa", "Nigeria",
  "Kormantine", "Ghana",
  "New Calabar", "Nigeria",
  "Bight of Benin", "Benin",
  "Popo", "Benin",
  "Cess", "Liberia",
  "Bance Island (Ben's Island)", "Sierra Leone",
  "Apammin", "Ghana",
  "Sierra Leone estuary", "Sierra Leone",
  "Senegambia and offshore Atlantic", "Senegal",
  "Portudal", "Senegal",
  "Mpinda", "Angola",
  "Little Popo", "Togo",
  "Lay", "Benin",
  "Goree", "Senegal",
  "Christiansborg", "Ghana",
  "Cape Mount", "Liberia",
  "Cape Mount (Cape Grand Mount)", "Liberia",
  "Cap Lopez", "Gabon",
  "Cameroons River", "Cameroon",
  "Bight of Biafra and Gulf of Guinea Islands", "Cameroon",
  "Pokesoe (Princes Town)", "Ghana",
  "Rio Forcados", "Nigeria", 
  "River del Rey", "Cameroon", 
  "Sao Tome", "Africa",
  "Sao Tome and Principe", "Africa",
  "Corisco", "Equatorial Guinea",
  "Cape Coast Castle", "Ghana",
  "St. Maarten", "Phillipsburg",
  "Rio de la Hacha", "Colombia"
)
```

# Functions ---------------------------------------------------------------

# Strips suffixes, whitespaces and translates LUT entries
```{r}
clean_places <- function(places, place_lut = place_lut) {
  suffixes <- c("port unspecified", "location unspecified", "colony unspecified")
  suffixes <- c(paste0(", ", suffixes), "province")
  suffixes <- paste0(suffixes, collapse = "|")
  places <- gsub(suffixes, "", places)
  places <- str_trim(places)
  m <- match(places, place_lut$replace)
  places[!is.na(m)] <- place_lut$by[m][!is.na(m)]
  places
}
```

# Cosine similarity
```{r}
cos_sim <- function(x) {
  x <- x / sqrt(rowSums(x * x))
  x <- x %*% t(x)
  x
}
```

# Cosine distance
```{r}
cos_dist <- function(x) {
  as.dist(1 - cos_sim(x))
}
```

# Data wrangling ----------------------------------------------------------

# Global wrangling --------------------------------------------------------

# Basically trying to convert placenames to country names.
```{r}

df <- slave_routes %>%
  # Clean placenames
  mutate(
    port_origin = clean_places(port_origin, place_lut = place_lut),
    port_arrival = clean_places(port_arrival, place_lut = place_lut),
    place_of_purchase = clean_places(place_of_purchase,place_lut = place_lut)
  ) %>%
  # Try to find countries by placename
  mutate(
    country_origin = ifelse(port_origin %in% cities$country.etc, port_origin, 
                            cities$country.etc[match(port_origin, cities$name)]),
    country_arrival = ifelse(port_arrival %in% cities$country.etc, port_arrival, 
                             cities$country.etc[match(port_arrival, cities$name)]),
    country_purchase = ifelse(place_of_purchase %in% cities$country.etc, place_of_purchase,
                              cities$country.etc[match(place_of_purchase, cities$name)])
  ) %>%
  # Replace NAs by place if country is not found
  mutate(country_purchase = ifelse(is.na(country_purchase), place_of_purchase, country_purchase)) %>%
  mutate(country_arrival = ifelse(is.na(country_arrival), port_arrival, country_arrival)) %>%
  arrange(-n_slaves_arrived)
```

# Netherlands wrangling ---------------------------------------------------

# Define a few patterns to counter overplotting

# Arrival patterns
```{r}
carribean_patterns <- c("Caribbean", "Barbados", "Jamaica", "Haiti", 
                        "Saint Kitts", "US Virgin", "Martinique", "Montserrat",
                        "Tobago", "Saint Lucia", "Guadeloupe")

carribean_patterns <- paste0(carribean_patterns, collapse = "|")
```

# Purchase pattern
```{r}
broad_pattern <- c("Windward", "Africa") # Should catch Windward + Ivory + Gold & Africa, West Central Africa
broad_pattern <- paste0(broad_pattern, collapse = "|")
```

# Wrangle data for the Netherlands
```{r}
netherlands <- df %>%
  filter(country_origin == "Netherlands" | country_arrival == 'Netherlands') %>%
  mutate(country_arrival = ifelse(grepl(carribean_patterns, country_arrival), "Other Carribean", country_arrival)) %>%
  mutate(country_arrival = ifelse(grepl("Guiana", country_arrival), "Guyana", country_arrival)) %>%
  mutate(country_arrival = ifelse(grepl("unspecified", country_arrival), "Unspecified", country_arrival)) %>%
  mutate(country_arrival = ifelse(grepl("Africa|Canary|Ghana", country_arrival), "Africa", country_arrival)) %>%
  mutate(country_arrival = ifelse(grepl("Americas", country_arrival), "Other Americas", country_arrival)) %>%
  mutate(country_purchase = ifelse(grepl(broad_pattern, country_purchase), "Broader\nAfrican Regions", country_purchase))
```

# Summarise the Netherlands
```{r}
nl_countries <- netherlands %>%
  group_by(country_purchase, country_arrival) %>%
  summarise(slaves = sum(n_slaves_arrived, na.rm = TRUE),
            voyages = length(unique(voyage_id))) %>%
  drop_na() %>%
  filter(slaves > 0) %>%
  arrange(-slaves)
```


# Parallel sets plot ------------------------------------------------------
```{r}
nl_plot <- gather_set_data(nl_countries, 1:2)

```

```{r}
g <- ggplot(nl_plot, aes(x, id = id, split = y, value = slaves)) +
  geom_parallel_sets(aes(fill = country_purchase), sep = 0.1) +
  geom_parallel_sets_labels(colour = 'black', angle = 0, sep = 0.1) +
  geom_parallel_sets_axes(axis.width = 0.1, sep = 0.1, colour = "white", fill = "white") +
  scale_x_discrete(limits = c("country_purchase", "country_arrival"))

label_dat <- layer_data(g, 2)
axis_dat <- layer_data(g, 3) %>%
  filter(!duplicated(cbind(label, xmin)))

country_exchange <- ggplot(nl_plot, aes(x, id = id, split = y, value = slaves)) +
  geom_parallel_sets(aes(fill = country_purchase, colour = country_purchase), sep = 0.1) +
  geom_rect(data = filter(axis_dat, xmin == 0.95),
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, 
                fill = label, colour = label),
            inherit.aes = FALSE) +
  geom_rect(data = filter(axis_dat, xmin == 1.95),
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            inherit.aes = FALSE, colour = "white", fill = "white") +
  geom_text(data = filter(label_dat, x == 1),
            aes(x, y, label = label),
            inherit.aes = FALSE, hjust = 1, nudge_x = -0.07,
            colour = "white") +
  geom_text(data = filter(label_dat, x == 2),
            aes(x, y, label = label),
            inherit.aes = FALSE, hjust = 0, nudge_x = 0.07,
            colour = "white") +
  annotate("rect", xmin = 2.4, xmax = 2.5, ymin = 0, ymax = 1e5, fill = "white", colour = "white") +
  annotate("text", x = 2.52, y = 5e4, label = "100.000\nArrived\nPeople", hjust = 0, colour = "white",
           size = 3) +
  geom_blank(data = data.frame(x = 2.6),aes(x = x), inherit.aes = FALSE) +
  scale_fill_discrete(aesthetics = c("fill", "colour"))+
                      #guide = guide_none()) +
  scale_x_discrete(limits = c("country_purchase", "country_arrival"),
                   labels = c("Purchase Country", "Arrival Country"),
                   expand = c(0,0.4,0,0.7), name = "") +
  scale_y_continuous(breaks = NULL, name = "") +
  labs(title = "Transatlantic slavery with ships departing from <i style='color:#FF7F00'>**Dutch**</i> ports",
       subtitle = paste0("Between the years ", paste0(range(netherlands$year_arrival), collapse = "-")),
       caption = "Data and statistics: https://slavevoyages.org/ \n
       City-country links: ?maps::world.cities") +
  theme(
    plot.background = element_rect(colour = NA, fill = "black"),
    panel.background = element_rect(colour = NA, fill = "grey10"),
    panel.grid.major = element_blank(),
    plot.title = element_markdown(colour = "white"),
    plot.subtitle = element_text(colour = "grey80"),
    plot.caption = element_text(colour = "white", lineheight = 0.5),
    axis.ticks = element_blank(),
    axis.text = element_text(colour = "white", size = rel(1))
  )
```


# Save plots --------------------------------------------------------------
```{r}
ggsave("2020_06_16_Slavery_Countries.png", plot = country_exchange,
       width = 8, height = 6, bg = "black")
```

